Legacy Data Confounds Genomics Studies

### 2019-07-12 LatexManuscript/legacy.tex
some comments and edits

%%% PREAMBLE 
\documentclass[9pt,article]{template}

@@

To investigate whether suspicious calls from the 1kGP are imputed into genotyping studies, we submitted genotype data for the first two chromosomes of the 1kGP genotype data to the Michigan Imputation Server.
We found that all of the variants associated with $Q$ were imputed back in the samples.
This suggests that the imputation reference panel still includes individuals with low $Q$, and the dubious variants will be imputed in individuals who most closely match the individuals with low $Q$.
These $Q$-associated variants could also compromise the imputation of nearby real-variants, however when considering the imputation scores of genotype data from Japanese individuals from Nagahama, there does not appear to be any impact on nearby variant imputation scores (Figure \ref{imputation_score}).

We searched the literature for any GWAS that might have reported these $Q$-associated variants as being significantly associated with some biological trait, even though there is no particular reason for these variants to be associated with phenotypes.
The NHGRI-EBI Catalog of published genome-wide association studies identified seventeen recent publications that had reported these variants as close to or above the genome-wide significant threshold (Table \ref{gwasTable}).

@@

All of these articles used a variety of strict quality filters, including Hardy-Weinberg equilibrium test, deviations in expected allele frequency and sequencing data quality thresholds.
They also removed rare alleles and alleles with high degrees of missingness.
Indeed, we expect a large number of $Q$-associated variants to be filtered out by some quality controls like the Hardy-Weinberg equilibrium test.
Even though the studies used state-of-the-art quality controls, the variants were imputed onto genotype data and reached genome wide significance for association with biological traits.
However, the fact that some of these variants in other studies are not removed and that the great majority of these variants are missing from higher quality datasets means that these $Q$-associated variants should be flagged for removal to avoid spurious association.

>>>>

%%% PREAMBLE 
\documentclass[9pt,lineno]{template}

@@

To investigate whether suspicious calls from the 1kGP are imputed into genotyping studies, we submitted genotype data for the first two chromosomes of the 1kGP genotype data to the Michigan Imputation Server.
We found that all of the variants associated with $Q$ were imputed back in the samples.
This suggests that the imputation reference panel still includes individuals with low $Q$, and the dubious variants will be imputed in individuals who most closely match the individuals with low $Q$.
These $Q$-associated variants could also compromise the imputation of nearby real variants because haplotype matching would be compromised. However when considering the imputation scores of genotype data from Japanese individuals from Nagahama, there does not appear to be any impact on nearby variant imputation scores (Figure \ref{imputation_score}).

We searched the literature for GWAS that have reported these $Q$-associated variants as being significantly associated with some biological trait, even though there is no particular reason for these variants to be associated with phenotypes.
The NHGRI-EBI Catalog of published genome-wide association studies identified seventeen recent publications that had reported these variants as close to or above the genome-wide significant threshold (Table \ref{gwasTable}).

@@

These articles used a variety of strict quality filters, including Hardy-Weinberg equilibrium test, deviations in expected allele frequency and sequencing data quality thresholds.
They also removed rare alleles and alleles with high degrees of missingness.
%Indeed, we expect a large number of $Q$-associated variants to be filtered out by some quality controls like the Hardy-Weinberg equilibrium test.
Despite state-of-the-art quality controls, the variants were imputed onto genotype data, reached genome wide significance for association with biological traits, and were reported to the GWAS catalogue.
However, the fact that some of these variants in other studies are not removed and that the great majority of these variants are missing from higher quality datasets means that these $Q$-associated variants should be flagged for removal to avoid spurious association. \sgcomment{I'm not sure what the previous sentence is saying}
